OCMu64: a solver for One-sided Crossing Minimization

### 2024-06-19 tex/description.tex
Change first lemma in description

\section{Definitions}
We use $<$ to compare vertices in $A$ in their fixed ordering. We generalize to weighted graphs
for the proof of \ref{lem:sdp}: a node $u \in B$ is taken to be a function $u: A \to \mathbb
R^{\geq 0}$, where weight 0 is an absent edge. We write $c(u, v) = \sum_{(a, b) \in A^2}
u(a)v(b)[a > b]$; in the unweighted case, this is the \emph{crossing number}, the number of
crossings between edges incident to $u$ and $v$ when $u$ is drawn before $v$.

For $X,Y\subseteq B$ we set $c(X,Y) = \sum_{x\in X}\sum_{y\in Y} c(x,y)$ for the cost of
ordering all vertices of $X$ before all vertices of $Y$. More generally, $c(X,Y,Z) = c(X, Y) +
c(X, Z) + c(Y, Z)$. We also consider the \emph{reduced cost} $r(X,Y) = c(X, Y) - c(Y, X)$,
which is negative when $X$ is better \emph{before} $Y$ and positive when $X$ is better
\emph{after} $Y$.

@@

\section{Methods}
\subsection{Reductions}
\subparagraph{Dominating pairs}
We give a much stronger version of \cref{lem:weak}. In fact, this
lemma is as strong as can be when only considering $N(u)$ and $N(v)$.

\begin{lemma}[Strong dominating pair]\label{lem:sdp}
  Write $n:= |N(u)|$, $m:=|N(v)|$, and consider both sets as sorted lists.
  When for all $i\in [n]$ we have
  $ N(u)_i \leq N(v)_{\lfloor i\cdot m/n \rfloor} $,
  then $u\be v$.
  When there is an $i$ such that the inequality is strict, then $u\b v$.
\end{lemma}

%Note that the condition is equivalent to the symmetric version: for all $j\in [m]$:
%$ U_{\lceil j\cdot n/m \rceil} \leq V_j $.

\begin{proof}
  Consider for each $u\in B$ the function $w_u: A\to \R_{\geq 0}$ that
  assigns \emph{weight} $1$ to neighbours of $u$ and weight $0$ otherwise.
  Let $W_u = \sum_{a\in A} w_u(A)$ be the total weight (degree) of $u$.
  Then for any $u$ and $v$, $c(u,v) = \sum_{a,b\in A, a>b} w_u(a) \cdot w_v(b)$.
  We can `shift' some weight $\delta$ of $w_u$ to the right from $a_1$ to $a_2>a_1$ by
  decreasing $w_u(a_1)\geq \delta$ by $\delta$ and increasing $w_u(a_2)$ by
  $\delta$.
  This increases the value of
  $c(u,v)$ by $\delta \cdot \sum_{a_1 < b \leq a_2} w_b(v)\geq 0$.

@@

\begin{remark}
  When the condition in \cref{lem:sdp} does not hold, there exists a multiset
  $\mathcal X\subseteq A$ such that $v, \mathcal X, u$ is the optimal order.
\end{remark}

>>>>

\section{Definitions}
We use $<$ to compare vertices in $A$ in their fixed ordering. We generalize to weighted graphs
for the proof of \ref{lem:sfp}: a node $u \in B$ is taken to be a function $u: A \to \mathbb
R^{\geq 0}$, where weight 0 is an absent edge. We write $W_u = \sum_{a \in A}u(a)$ for the
total weight of $u$, and set $\bar u = u/W_u$.

We write $c(u, v) = \sum_{(a, b) \in A^2}
u(a)v(b)[a > b]$; in the unweighted case, this is the \emph{crossing number}, the number of
crossings between edges incident to $u$ and $v$ when $u$ is drawn before $v$.  For
$X,Y\subseteq B$ we set $c(X,Y) = \sum_{x\in X}\sum_{y\in Y} c(x,y)$ for the cost of ordering
all vertices of $X$ before all vertices of $Y$. More generally, $c(X,Y,Z) = c(X, Y) + c(X, Z) +
c(Y, Z)$. We also consider the \emph{reduced cost} $r(X,Y) = c(X, Y) - c(Y, X)$, which is
negative when $X$ is better \emph{before} $Y$ and positive when $X$ is better \emph{after} $Y$.

@@

\section{Methods}
\subsection{Reductions}
\subparagraph{Fixed pairs}
We give a much stronger version of \cref{lem:weak}. In fact, this
lemma is as strong as can be when only considering $u, v$ themselves. We begin with this
abstract version, followed by a concrete corrolary.

\begin{lemma}[Strong fixed pair]\label{lem:sfp}
    Fix $u, v$. Suppose that for each $b \in A$ we have $\sum_{a \leq b} \bar u(a) \geq \sum_{a
    \leq b} \bar v(a)$. Then, for any $w$, we have $r(w, \bar v) \leq r(w, \bar u)$.
\end{lemma}

\begin{proof}
    If $\bar u \neq \bar v$, consider the least element $a_0 \in A$ such that $\bar u(a_0) \neq
    \bar v(a_0)$. We must have $\bar u(a_0) > \bar v(a_0)$. Consider the transformation of
    $\bar u$ which ``moves'' $\bar u(a_0) - \bar v(a_0)$ of weight from $a_0$ to its successor
    $a_1$, and call this transformed function $\bar u'$. Then
    \[
        r(w, \bar u') = r(w, \bar u) - ... \leq r(w, \bar u).
    \]
    Since $\bar v$ can be obtained from $\bar u$ by a finite sequence of such transformations,
    the inequality follows.
\end{proof}

% To do, fix everything from here.
\begin{proof}
  Consider for each $u\in B$ the function $w_u: A\to \R_{\geq 0}$ that
  assigns \emph{weight} $1$ to neighbours of $u$ and weight $0$ otherwise.
  Let $W_u = \sum_{a\in A} w_u(A)$ be the total weight (degree) of $u$.
  Then for any $u$ and $v$, $c(u,v) = \sum_{a,b\in A, a>b} w_u(a) \cdot w_v(b)$.
  We can `shift' some weight $\delta$ of $w_u$ to the right from $a_1$ to $a_2>a_1$ by
  decreasing $w_u(a_1)\geq \delta$ by $\delta$ and increasing $w_u(a_2)$ by
  $\delta$.
  This increases the value of
  $c(u,v)$ by $\delta \cdot \sum_{a_1 < b \leq a_2} w_b(v)\geq 0$.

@@

\begin{remark}
  When the condition in \cref{lem:sfp} does not hold, there exists a multiset
  $\mathcal X\subseteq A$ such that $v, \mathcal X, u$ is the optimal order.
\end{remark}
